Comparing Machine Learning algorithms for classifying cognitive states from fMRI data

### 2012-11-09 article/article.tex
barry

%% Abstract section.
\abstract{Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain. This technique is typically used to determine which areas of the brain are involved in performing a certain task. Using this knowledge, we may also be able to determine what task was performed by looking at fMRI data. In this paper, we will discuss the use of machine learning algorithms for the classification of fMRI data. By using classifiers we will try to predict the cognitive state of subjects, based on their brain activity during that time.
} % end of abstract

@@

ROI mapping. As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field. To reduce the number of features, average voxel values within this regions are taken as input features. As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not utilized.

Principal Component Analysis (PCA).

@@

\section{Experimental Results}
\label{sec:results}

As a comparison measure, we used the performance of the classifiers for predicting the cognitive state that the subject was in. Table~\ref{tab:results} illustrates the overall performance of the three classifiers. Rows indicate various feature selection methods, where \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, and \emph{SD} for \emph{Standard Deviation}. For instance, \emph{Norm-M-SD ROIs} denotes that normalisation was used in the voxels belonging to the selected ROIs over the mean and standard deviation.

\begin{table*}[htpb]
\centering
{\small
\begin{tabular}{c|c|c|c|c}
& \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
PCA & 61.46 &  - & - & - \\
M ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
Norm-M ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
Norm-M-SD ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
\end{tabular}}
\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure.}
\label{tab:results}
\end{table*}

Columns depict the used classifiers---Naive Bayes, Linear SVM, and Multivariate. For the classifiers depicted with *, we used different structure for the data that we fit to them, as described in Section~\ref{}. The table reveals the following general results:

\begin{itemize}
	\item Classifiers performance do not differ significantly between them. (explain)
	\item Classifiers performance differ significantly from 50\%.
\end{itemize}

Moreover, it is observed that the Multivariate case has the best performance between all classifiers (why?). When PCA is used for feature extraction, the performance of Naive Bayes is lower than when other feature selection methods are used (why?).

We can also distinguish the fact that when different structure is used for the fitting data, the performance increases, though not significantly. This is observed in the case of the Naive Bayes classifier. When different structure is used, its performance increases approximately 4\%.

%mention confidence interval (how much percent sure we are that the actual accuracy of the classifier is between these regions)

>>>>

%% Abstract section.
\abstract{
Functional magnetic resonance imaging (fMRI) is a brain mapping technique that allows researchers to measure activity in a subject's brain. This technique is typically used to determine which areas of the brain are involved in performing a certain task. Using this knowledge, we may also be able to determine what task was performed by looking at fMRI data. In this paper, we will discuss the use of machine learning algorithms for the classification of fMRI data. By using classifiers we will try to predict the cognitive state of subjects, based on their brain activity during that time.
} % end of abstract

@@

ROI mapping. ROI mapping is a form of feature selection, wherein only a subset of the d-dimensional input space is used. As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field. To reduce the number of features, average voxel values within this regions are taken as input features. As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis.

Principal Component Analysis (PCA). As feature extraction method, we applied PCA. In between subject classification, the input vectors of the PCA should ideally be the standardized Talairach coordinates. As this coordinate system appeared to be inaccurate for this dataset, we were forced to use the voxel indices. Because brain sizes and shapes differ per subject, it is uncertain whether the voxel indices map well onto each other. As a result, the accuracy of the PCA is reduced.

@@

\section{Experimental Results}
\label{sec:results}
- raw results per algorithm
